---
title: "ENVS 193DS Homework 3"
author: "Brooke Ryan"
format: docx
editor: visual
---

# link to forked repository:

https://github.com/Brookepryan/ryan_brooke_homework-03.git

# reading in packages/data

```{r}
#suppressing messages/warnings
suppressMessages({
  suppressWarnings({
# general use
library(tidyverse)
library(readxl)
library(here)
library(janitor)
library(dplyr)
library(ggplot2)
    
# visualizing pairs
library(GGally)

# model selection
library(MuMIn)

# model predictions
library(ggeffects)

# model tables
library(gtsummary)
library(flextable)
library(modelsummary)

drought_exp <- read_xlsx(path = here("data", 
                                     "Valliere_etal_EcoApps_Data.xlsx"),
                         sheet = "First Harvest")
                        
# quick look at data 
str(drought_exp)
class(drought_exp)
  })
})

```

```{r}
#suppressing messages/warnings
suppressMessages({
  suppressWarnings({
    
# cleaning
drought_exp_clean <- drought_exp %>% 
  clean_names() %>% # nicer column names
  mutate(species_name = case_when( # adding column with species scientific names
    species == "ENCCAL" ~ "Encelia californica", # bush sunflower
    species == "ESCCAL" ~ "Eschscholzia californica", # California poppy
    species == "PENCEN" ~ "Penstemon centranthifolius", # Scarlet bugler
    species == "GRICAM" ~ "Grindelia camporum", # great valley gumweed
    species == "SALLEU" ~ "Salvia leucophylla", # Purple sage
    species == "STIPUL" ~ "Nasella pulchra", # Purple needlegrass
    species == "LOTSCO" ~ "Acmispon glaber" # deerweed
  )) %>% 
  relocate(species_name, .after = species) %>% # moving species_name column after species
  mutate(water_treatment = case_when( # adding column with full treatment names
    water == "WW" ~ "Well watered",
    water == "DS" ~ "Drought stressed"
  )) %>% 
  relocate(water_treatment, .after = water) # moving water_treatment column after water

 })
})
```

# a. Make a table or list of all the models from class and the last one you constructed on your own. Write a caption for your table

## 0. Null model

```{r}
model0 <- lm(total_g ~ 1, # formula
             data = drought_exp_clean) # data frame
```

## 1. total biomass as a function of SLA, water treatment, and species

```{r}
# saturated model
model1 <- lm(total_g ~ sla + water_treatment + species_name,
             data = drought_exp_clean)

par(mfrow = c(2, 2))
plot(model1)
# you might get a warning when you run this code - that is ok!
#diagnostics for sat model look good
```

## 2. total biomass as a function of SLA and water treatment

```{r}
model2 <- lm(total_g ~ sla + water_treatment,
             data = drought_exp_clean)

plot(model2)
```

**look at diagnostics later**

## 3. total biomass as a function of SLA and species

```{r}
model3 <- lm(total_g ~ sla + species_name,
             data = drought_exp_clean)

plot(model3)
```

## 4. total biomass as a function of water treatment and species

```{r}
model4 <- lm(total_g ~ water_treatment + species_name,
             data = drought_exp_clean)

plot(model4)
```

```{r}
summary(model4)
```

# Making a table

```{r, fig.with = 9, fig.height = 12}

modelsummary::modelsummary( # this function takes a list of models
  list( 
    "null" = model0, # "model name" = model object
    "model 1" = model1,
    "model 2" = model2,
    "model 3" = model3,
    "model 4" = model4)) 

```

# Caption:

### Table 1: Comparison of Linear Regression Models for Predicting Total Growth

This table shows the coefficients, standard errors, and performance metrics for four linear regression models that predict total growth biomass (`total_g`) based on water treatment and species name. Model 4, which includes both predictors, shows the best performance with the highest R² (0.754) and lowest AIC (-159.2), indicating the most reliable fit for the data.

e.  Statistical Methods

To examine the influence of specific leaf area (SLA), species, and water treatment on total plant biomass of plants species native to southern California, I conducted a series of linear regression models. I used a null model to establish a baselines, and then four subsequent models comparing different combinations of predictors (SLA, species, water treatment). To determine the model that best described total biomass, I used Akaike Information Criterion (AIC) and R² values. I decided that the model that best described variation in plant biomass is model 4, because it had the highest R² of 0.754 and lowest AIC of -159.2. To evaluate linear model assumptions, I visually inspected diagnostic plots for normality, homoscedasticity, linearity, and outliers.

# c. Visualization of model 4, the best predictor.

{r, fig.width = 8, fig.height = 5}

suppressMessages({ \# suppress warnings and messages suppressWarnings({

model4_preds \<- ggpredict(model4, \# getting predictions terms = c("water_treatment", "species_name"))

# plotting predictions + underlying data

ggplot(model4_preds, aes(x = x, y = predicted, color = x)) + \# plotting predictions geom_jitter(data = drought_exp_clean, aes(x = water_treatment, y = total_g, color = water_treatment), width = 0.2, alpha = 0.1) + \# plotting underlying data geom_line() + geom_point() + geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) + \# representing st error theme_classic() + labs(title = "Predicted Impact of Water Treatment on Plant Biomass Across Different Species", \# adding and changing titles x = "Water Treatment", y = "Biomass") + scale_color_manual(name = "Water Treatment", values = c("Well watered" = "seagreen2", "Drought stressed" = "turquoise3")) + \# editing colors theme(panel.grid = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1, size = 7.5), #changing orientation and size of x-axis text legend.position = "none") + \# remove legend facet_wrap(\~ group) \# facet by species name }) })

d.  Figure 1. "Predicted Impact of Water Treatment on Plant Biomass Across Different Species" This visualization compares predicted biomass of various plant species under two water treatment conditions: well watered and drought stressed. The error bars represent the confidence intervals of the predicted values while he underlying data is represented through jitter. Data source: Valliere J. M., J. Zhang, M. R. Sharifi, and P. W.Rundel. 2019. Can we condition native plants to increase drought tolerance and improve restoration success? Ecological Applications 29(3): e01863.10.1002/eap.1863

e.  Results Section

In conclusion, water treatment condition and species type are the most effective predictors of the total mass of native plants in Southern California, as demonstrated by the linear model with the highest R² (0.754) and lowest AIC (-159.2). On average plants in well watered treatment are 0.11695 ± 0.01733g greater than plants in drought stressed treatment. On average, Scarelet Bulgers (Penstemon centranthifolius) were the smallest and Great Valley Gumweed (Grindelia camporum) were the largest. Purple Needlegrass (Nasella pulchra) were, on average, the same size as California Poppy (Eschscholzia californica) were the same

# Problem 2. Affective visualization

## a. What would an affective visualization look like for my personal data?

An affective visualization of my data would include as many variables as possible. For this, I would probably be visualizing each observation individually and grouping them by day. I want to represent mg, form, pleasure/purpose, and hours of sleep at the very least. I think a drawing similar to the "dear data" project would be most fitting to accomplish this.

## b. Visualization sketch:

## c. Visualization draft:

## d. Write an artist statement.

An artist statement gives the audience context to understand your work. Write 4-5 sentences to address:

the content of your piece (what are you showing?) the influences (what did techniques/artists/etc. did you find influential in creating your work?) the form of your work (written code, watercolor, oil painting, etc.) your process (how did you create your work?)

My piece represents the amount of caffeine I consume per day in milligrams while representing variables such as the amount if sleep I got the night before, the form of caffeine I am consuming, and the reason I consumed the caffeine. I was largely influenced from the "Dear Data" project by Giorgia Lupi and Stefanie Posavec. My work is pen and colored pencil on paper, and I had to do many rough drafts in order to perfect it. I tried out various types of visual representations for each form of caffeine, and the amount of mg it represented, until I found a method that seemed clear.

# Problem 3. Statistical critique

## a. Revisit and summarize

The authors are using a one way analysis of variance (ANOVA) and two-way ANOVA both followed by Tukey’s post hoc test to address the question of better understand the shell calcification responses of P. fucata to ocean acidification. Specifically, they are testing the effects of lowered seawater pH on calcium content and shell hardness.

## b. Visual clarity

These visualizations are moderately clear representations to those who are skimming the article or do not know much about marine animal calcification. Some aspects could have been more defined, such as the lack of titles and the y-axis labeled “IcT”, considering I had never heard of that measuring unit so could not immediately understand the plot at first glance. The plots did show summary measurements including mean and standard error. The plots do not show underlying data.

## c. Aesthetic clarity

The plots are not cluttered at all, I would say that they are almost too bland and uninteresting. The data: ink ratio is good considering they are bar plots.

## d. Recommendations

I recommend doing a scatter plot to represent each variable instead of a box plot. This way, they could include the underlying data because it is a respectable amount of observations to have in a jitterplot, as well as an errorbarr to provide a clearer representation of mean and standard error. Also, I recommend specifying IcT as Growth Global Index of length and weight on the y-axis, and even possibly including the equation (IcT = (Ln At - Ln A0)/t\]100) somewhere on the plot, or at least in the caption. Lastly, I recommend including a title for each of their plots, it is confusing to read without one.
